Relationships in the extremes and their influence on competition and coexistence

### 2022-01-27 paper/Paper.Rnw
small things and some useful comments

%G2 
Fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
of the contributions to coexistence that come from ATAs compared to other mechanisms), Fig.
\ref{fig:fig3} shows that the ATA term of (\ref{eq:full_decomp}) often, but not always, has 
magnitude comparable to the magnitudes of other terms of the extended MCT decomposition. 
Magnitudes of $\Delta_i^{[EC]}$ tended to be larger, relative to the magnitudes of the 
other terms of (\ref{eq:full_decomp}), for smaller-magnitude differences between $\mu_1$ and 
$\mu_2$, i.e., for closer-to-equal competitors. 
In the $\mu_1=\mu_2$ case (panels A-D), for which the two species were equal 
competitors, storage effects (blue plus red
lines on Fig. \ref{fig:fig3}) were the only means by which coexistence could occur, and ATA effects
tended to be negative, weakening total storage effects.

%G3
Fulfilling goal G3 of the Introduction (to explore, using the lottery model, whether species
can coexist with ATAs, but not without, or vice versa), our results show that ATA can, indeed, sometimes
make the difference between coexistence and exclusion; both ``ATA rescue'' and ``ATA
exclusion'' (see Methods) are possible.
For some of the panels on Fig. \ref{fig:fig3}, the IGR line falls between the ATA
line and the $x$ axis for some values of $\sigma$, e.g., see panel H. 
Such cases correspond to parameter combinations
for which the presence or absence of ATAs determine whether the species coexist or not. In the 
case of panel H, for values of $\sigma$ around $4$, 
species 1 is excluded with ATAs but the two species can coexist without ATAs. 
Fig. \ref{fig:fig4} highlights additional examples. Orange lines on Fig. \ref{fig:fig4}
represent the IGR when the effects of ATAs are included, and blue lines are the IGR when
the effects of ATAs are removed. These two lines frequently crossed the $x$ axis in 
different places, showing that both ATA rescue space and ATA exclusion space can be non-empty
subsets of our trait-difference space (see Methods). For the parameters we considered, 
both ATA rescue and exclusion were more common for larger values of $\sigma$.
ATA rescue appeared more likely for smaller $\delta$, whereas ATA exclusion appeared
more common for larger $\delta$. Although bands of ATA rescue or exclusion on Fig.
\ref{fig:fig4} appear narrow, they nevertheless do occur, which means ATAs can mean the 
difference between coexistence and extirpation of one species in some cases.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig4.pdf}
\caption{ATAs can sometimes make the difference between species coexistence and exclusion of the weaker
competitor; both ``ATA rescue'' and ``ATA exclusion'' (see Methods) are possible. 
All panels show the IGR, the ATA term of the extended MCT [$\Delta_i^{[E || C]}$ in equation 
(\ref{eq:full_decomp})], and the difference of these quantities. 
For each panel, $\delta$ and $\sigma$ are fixed at values specified in the 
margins, and the three quantities are plotted against $\mu_1-\mu_2$ to isolate points in the
``trait-difference space'' corresponding to ATA rescue 
($\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$) 
or ATA exclusion ($\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$; 
see Methods). Values of $\mu_1-\mu_2$ for which the IGR is positive but 
$\text{IGR}-\Delta_i^{[EC]}$ is negative correspond to ATA rescue (shaded yellow), whereas
values of $\mu_1-\mu_2$ for which the IGR is negative but $\text{IGR}-\Delta_i^{[EC]}$ is positive 
correspond to ATA exclusion (shaded violet). Top panels also include domains
of ATA rescue or exclusion, but they were too narrow to be highlighted by shading and are
probably of minimal practical importance. 
Standard errors of plotted quantities were never greater than \Sexpr{round(se4,5)}, 
so are not displayed. The dashed red line on panel L corresponds to the parameters
used for Fig. \ref{fig:pedagog}. This plot uses $q_{ij}=1$. An analogous plot was generated 
using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$, but it looked exactly the same, by mathematical necessity (see SI section \ref{SI-sect:F4andqij}).}\label{fig:fig4}
\end{figure}

%***DAN: Jasmin, some changes are needed to the figure itself. Please make the text in the legend bigger.
%It looks as though a blue line is for ATA effects, and a red line is for the effects of correlation per se
%(that's what it says in the legend),
%but that differs from how you did it in the other figs. Or perhaps the legend is just wrong. Please check and
%fix so as to be consistent with the rest of the figures in which color represents which quantity and which
%notation represents which quantity. Also, please change panel labels to be of the format "(a)" instead of 
%"A". Make the panel labels bigger, as well as the axis numbers and labels. 
\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
\caption{The contribution of ATAs to species coexistence in the diatom system of \cite{Descamps_2005} when 
\emph{F. crotonensis} is the invader is meaningfully large compared to 
other components of the MCT decomposition. All panels show the IGR and the decomposition of
the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
%***DAN: This is for one species invading the other, right? Need to specify that, and say which is 
%invading which, and, at the end of the 
%caption indicate that the reverse invasion scenario is in sup mat and cite the corresponding figure there.
%***JASMIN: Done - ish. Need to make sup mat section 
For each panel, 
two of the three parameters defining the temperature oscillation (i.e., the amplitude, $a$; period, $P$; 
and mean, $\theta_0$; see Methods) are fixed at the original values used by 
\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, $\theta_0=18^\circ$C) and 
%***DAN: I added units for a  and \theta_0 here, please check these are correct 
%***JASMIN: Done 
the IGR and components of the MCT decomposition
are then plotted against the other parameter. The original value of the varying parameter 
is indicated by the vertical dashed magenta lines. ATA contributions 
had magnitude comparable to other components, 
%***DAN: Make sure this is still true after the correct version of C is put in, and after we
%resolve the potential problem with the wiggles on the plot. Reword if not.
indicating that ATAs can contribute meaningfully
to diatom coexistence in this system. In some instances, the IGR line (orange) falls 
between the ATA line (red)
and the horizontal axis, indicating that, for those parameter values, ATAs make the difference between species
coexistence and exclusion of the invader, \emph{F. crotonensis}. 
%***DAN: Not the weaker competitor, actually, right, but rather whichever competitor was being considered as the invader
%for these plots? If I'm right about that then please reword accordingly. 
%***JASMIN: Right, Done. 
Standard errors of plotted quantities ?.
%***DAN: Don't forget to fill in the standard error stuff
%***DAN: Here is where you add a statement citing another figure in the sup mat that takes the other
%species as the invader.
%***JASMIN: Done
Coexistence in this system and within 
our defined parameters only depends on the IGR of \emph{F. crotonensis} so we only 
show this case; though, an analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
\emph{F. crotonensis} is the resident.}  \label{fig:fig5}
\end{figure}

>>>>

%G2 
Fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
of the contributions to coexistence that come from ATAs compared to other mechanisms), Fig.
\ref{fig:fig3} shows that the ATA term of (\ref{eq:full_decomp}) often, but not always, has 
magnitude comparable to the magnitudes of other terms of the extended MCT decomposition. 
Magnitudes of $\Delta_i^{[EC]}$ tended to be larger, relative to the magnitudes of the 
other terms of (\ref{eq:full_decomp}), for smaller-magnitude differences between $\mu_1$ and 
$\mu_2$, i.e., for closer-to-equal competitors. 
In the $\mu_1=\mu_2$ case (panels a-d), for which the two species were equal 
competitors, storage effects (blue plus red
lines on Fig. \ref{fig:fig3}) were the only means by which coexistence could occur, and ATA effects
tended to be negative, weakening total storage effects.

%G3
Fulfilling goal G3 of the Introduction (to explore, using the lottery model, whether species
can coexist with ATAs, but not without, or vice versa), our results show that ATAs can, indeed, sometimes
make the difference between coexistence and exclusion: both ``ATA rescue'' and ``ATA
exclusion'' (see Methods) are possible.
For some of the panels on Fig. \ref{fig:fig3}, the IGR line falls between the ATA
line and the $x$ axis for some values of $\sigma$, e.g., see panel h. 
Such cases correspond to parameter combinations
for which the presence or absence of ATAs determine whether the species coexist or not. In the 
case of panel h, for values of $\sigma$ around $4$, 
species 1 is excluded with ATAs but the two species can coexist without ATAs. 
Fig. \ref{fig:fig4} highlights additional examples. Orange lines on Fig. \ref{fig:fig4}
represent the IGR when the effects of ATAs are included, and blue lines are the IGR when
the effects of ATAs are removed. These two lines frequently crossed the $x$ axis in 
different places, showing that both ATA rescue space and ATA exclusion space can be non-empty
subsets of our trait-difference space (see Methods). For the parameters we considered, 
both ATA rescue and exclusion were more common for larger values of $\sigma$.
ATA rescue appeared more likely for smaller $\delta$, whereas ATA exclusion appeared
more common for larger $\delta$. Although some bands of ATA rescue or exclusion on Fig.
\ref{fig:fig4} appear narrow, they nevertheless do occur, which means ATAs can mean the 
difference between coexistence and extirpation of one species in some cases.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig4.pdf}
\caption{ATAs can sometimes make the difference between species coexistence and exclusion of the weaker
competitor: both ``ATA rescue'' and ``ATA exclusion'' (see Methods) are possible. 
All panels show the IGR, the ATA term of the extended MCT [$\Delta_i^{[EC]}$ in equation 
(\ref{eq:full_decomp})], and the difference of these quantities. 
For each panel, $\delta$ and $\sigma$ are fixed at values specified in the 
margins, and the three quantities are plotted against $\mu_1-\mu_2$ to isolate points in the
``trait-difference space'' corresponding to ATA rescue 
($\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$) 
or ATA exclusion ($\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$; 
see Methods). Values of $\mu_1-\mu_2$ for which the IGR is positive but 
$\text{IGR}-\Delta_i^{[EC]}$ is negative correspond to ATA rescue (shaded yellow), whereas
values of $\mu_1-\mu_2$ for which the IGR is negative but $\text{IGR}-\Delta_i^{[EC]}$ is positive 
correspond to ATA exclusion (shaded violet). Top panels also include domains
of ATA rescue or exclusion, but they were too narrow to be highlighted by shading and are
probably of minimal practical importance. 
Standard errors of plotted quantities were never greater than \Sexpr{round(se4,5)}, 
so are not displayed. The dashed red line on panel (l) corresponds to the parameters
used for Fig. \ref{fig:pedagog}. This plot uses $q_{ij}=1$. An analogous plot was generated 
using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$, but it looked exactly the same, by mathematical necessity (see SI section \ref{SI-sect:F4andqij}).}\label{fig:fig4}
\end{figure}

%***DAN: Jasmin, some changes are needed to the figure itself. Please make the text in the legend bigger.
%It looks as though a blue line is for ATA effects, and a red line is for the effects of correlation per se
%(that's what it says in the legend),
%but that differs from how you did it in the other figs. Or perhaps the legend is just wrong. Please check and
%fix so as to be consistent with the rest of the figures in which color represents which quantity and which
%notation represents which quantity. Also, please change panel labels to be of the format "(a)" instead of 
%"A". Make the panel labels bigger, as well as the axis numbers and labels. 
%***DAN: If I interpret correctly that the blue lines on Fig. 5 are the ATA contribution (needs to be 
%changed because you use a red line on other figs), there are large areas on panels a and c where the 
%IGR is between the ATA-contribution line and then x axis. Those are areas of ATA rescue (though you only
%know this after looking at the analogous plot for the other species). I think in 
%the next version of the figure you should consider highlighting them in a similar way to how you did 
%in Fig. 4. Those are a couple nice wide zones of ATA rescue! Let me know if you want to talk about this.
\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
\caption{The contribution of ATAs to species coexistence in the diatom system of \cite{Descamps_2005} when 
\emph{F. crotonensis} is the invader is meaningfully large compared to 
other components of the MCT decomposition. All panels show the IGR and the decomposition of
the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
%***DAN: This is for one species invading the other, right? Need to specify that, and say which is 
%invading which, and, at the end of the 
%caption indicate that the reverse invasion scenario is in sup mat and cite the corresponding figure there.
%***JASMIN: Done - ish. Need to make sup mat section 
For each panel, 
two of the three parameters defining the temperature oscillation (i.e., the amplitude, $a$; period, $P$; 
and mean, $\theta_0$; see Methods) are fixed at the original values used by 
\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, $\theta_0=18^\circ$C) and 
%***DAN: I added units for a  and \theta_0 here, please check these are correct 
%***JASMIN: Done 
the IGR and components of the MCT decomposition
are then plotted against the other parameter. The original value of the varying parameter 
is indicated by the vertical dashed magenta lines. ATA contributions 
had magnitude comparable to other components, 
%***DAN: Make sure this is still true after the correct version of C is put in, and after we
%resolve the potential problem with the wiggles on the plot. Reword if not.
indicating that ATAs can contribute meaningfully
to diatom coexistence in this system. In some instances, the IGR line (orange) falls 
between the ATA line (red)
and the horizontal axis, indicating that, for those parameter values, ATAs make the difference between species
coexistence and exclusion of the invader, \emph{F. crotonensis}. 
%***DAN: Not the weaker competitor, actually, right, but rather whichever competitor was being considered as the invader
%for these plots? If I'm right about that then please reword accordingly. 
%***JASMIN: Right, Done. 
Standard errors of plotted quantities ?.
%***DAN: Don't forget to fill in the standard error stuff
%***DAN: Here is where you add a statement citing another figure in the sup mat that takes the other
%species as the invader.
%***JASMIN: Done
Coexistence in this system and within 
our defined parameters only depends on the IGR of \emph{F. crotonensis} so we only 
show this case; though, an analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
\emph{F. crotonensis} is the resident.}  \label{fig:fig5}
\end{figure}
